---
title: "R Notebook"
output: html_notebook
---

# Answering the following business questions 

# What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)
# What was the average age of people who are going out trick or treating?
# What was the average age of people who are not going trick or treating?
# For each of joy, despair and meh, which candy bar received the most of these ratings?
# How many people rated Starburst as despair?

```{r}
#loading necessary packages
library(tidyverse)
```

```{r}
#loading the clean data
candy_clean <- read_csv("../clean_data/candy_clean.csv")
candy_clean
```

```{r}
# What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)
candy_clean %>% 
  drop_na(rating) %>% 
  count(rating)
```
# The number of ratings across the three years is 339497 (DESPAIR), 293891 (JOY), 88518 (MEH)


```{r}
#What was the average age of people who are going out trick or treating?
candy_clean %>% 
  filter(trick_or_treat == "Yes") %>% 
  drop_na(age) %>% 
  summarise(mean_age = mean(age, na.rm = TRUE))
```

# The mean age of people going trick or treating is 35 years old 

```{r}
#What was the average age of people who are not going trick or treating?
candy_clean %>% 
  filter(trick_or_treat == "No") %>% 
  drop_na(age) %>% 
  summarise(mean_age = mean(age, na.rm = TRUE))
```
# The mean age of people not going trick or treating is 39 years old 


```{r}
#For each of joy, despair and meh, which candy bar received the most of these ratings?
candy_clean %>% 
  filter(rating == "JOY") %>% 
  count(candy_type) %>%
  arrange(n)
```
# Reese peanut butter cups is the candy with most joy ratings 

```{r}
#For each of joy, despair and meh, which candy bar received the most of these ratings?
candy_clean %>% 
  filter(rating == "MEH") %>% 
  count(candy_type) %>%
  arrange(n)
```

# Lollipops is the candy with most meh ratings

```{r}
#For each of joy, despair and meh, which candy bar received the most of these ratings?
candy_clean %>% 
  filter(rating == "DESPAIR") %>% 
  count(candy_type) %>%
  arrange(n)
```
# Broken glow sticks is the candy with most despair ratings



```{r}
#How many people rated Starburst as despair?
candy_clean %>% 
  group_by(candy_type, rating) %>% 
  filter(candy_type == "starburst", 
         rating == "DESPAIR") %>% 
  count(rating)
```

# 1990 people rated starburst as despair

> For the next three questions, count despair as -1, joy as +1, meh as 0

# What was the most popular candy bar by this rating system for each gender in the dataset ?
# What was the most popular candy bar in each year?
# What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?

```{r}
# add column to dataset with scores for each rating
candy_scored <- candy_clean %>% 
  mutate(score = case_when(
    rating == "JOY" ~ 1,
    rating == "DESPAIR" ~ -1,
    TRUE ~ 0
  ))

```

```{r}
#What was the most popular candy bar by this rating system for each gender in the dataset ?
candy_scored %>% 
  filter(!is.na(gender)) %>% 
group_by(gender, candy_type) %>% 
  summarise(total_score = sum(score)) %>% 
  slice_max(total_score)
```

#All respondents rated any full sized candy bar as the most popular 


```{r}
# What was the most popular candy bar in each year?
candy_scored %>% 
  group_by(timestamp, candy_type) %>% 
  summarise(total_score = sum(score)) %>% 
  slice_max(total_score)
```

#The most popular candy in 2015 and 2017 was Reese's Peanut Butter Cups, in 2016 the most popular candy was Kit Kat.

```{r}
# What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?
candy_scored %>%
  filter(!is.na(country)) %>% 
  group_by(country, candy_type) %>% 
  summarise(total_score = sum(score),
            num_votes = n()) %>%
  arrange(desc(total_score), num_votes) %>%
  slice_max(total_score, with_ties = F) %>% 
  select(country, candy_type)
```
#The most popular candy in the USA, UK AND Canada is "Any full sized candy bar". 