---
title: "R Notebook"
output: html_notebook
---

# Answering the following business questions 

# What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)
# What was the average age of people who are going out trick or treating?
# What was the average age of people who are not going trick or treating?
# For each of joy, despair and meh, which candy bar received the most of these ratings?
# How many people rated Starburst as despair?

```{r}
#loading necessary packages
library(tidyverse)
```

```{r}
#loading the clean data
candy_clean <- read_csv("../clean_data/candy_clean.csv")
candy_clean
```

```{r}
# What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)
candy_clean %>% 
  drop_na(rating) %>% 
  count(rating)
```
# The number of ratings across the three years is 339497 (DESPAIR), 293891 (JOY), 88518 (MEH)


```{r}
#What was the average age of people who are going out trick or treating?
candy_clean %>% 
  filter(trick_or_treat == "Yes") %>% 
  drop_na(age) %>% 
  summarise(mean_age = mean(age, na.rm = TRUE))
```

# The mean age of people going trick or treating is 35 years old 

```{r}
#What was the average age of people who are not going trick or treating?
candy_clean %>% 
  filter(trick_or_treat == "No") %>% 
  drop_na(age) %>% 
  summarise(mean_age = mean(age, na.rm = TRUE))
```
# The mean age of people not going trick or treating is 39 years old 


```{r}
#For each of joy, despair and meh, which candy bar received the most of these ratings?
candy_clean %>% 
  filter(rating == "JOY") %>% 
  count(candy_type) %>%
  arrange(n)
```
# Reese peanut butter cups is the candy with most joy ratings 

```{r}
#For each of joy, despair and meh, which candy bar received the most of these ratings?
candy_clean %>% 
  filter(rating == "MEH") %>% 
  count(candy_type) %>%
  arrange(n)
```

# Lollipops is the candy with most meh ratings

```{r}
#For each of joy, despair and meh, which candy bar received the most of these ratings?
candy_clean %>% 
  filter(rating == "DESPAIR") %>% 
  count(candy_type) %>%
  arrange(n)
```
# Broken glow sticks is the candy with most despair ratings



```{r}
#How many people rated Starburst as despair?
candy_clean %>% 
  group_by(candy_type, rating) %>% 
  filter(candy_type == "starburst", 
         rating == "DESPAIR") %>% 
  count(rating)
```

# 1990 people rated starburst as despair

> For the next three questions, count despair as -1, joy as +1, meh as 0

# What was the most popular candy bar by this rating system for each gender in the dataset ?
# What was the most popular candy bar in each year?
# What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?

```{r}
candy_clean <- candy_clean %>% 
 mutate(candy_rating = case_when (rating == "DESPAIR" ~ -1,
                                              rating == "JOY" ~ 1,
                                              TRUE ~ 0))
candy_clean
```
```{r}
candy_clean %>% 
  count(candy_rating)
```

```{r}
#What was the most popular candy bar by this rating system for each gender in the dataset ?
candy_clean %>% 
filter(gender == "Male") %>% 
  count(rating)
```
```{r}
candy_clean %>% 
  filter(gender == "male") %>% 
  group_by(rating) %>% 
  summarise(new_rating_total = sum(rating)) %>% 
  arrange(desc(new_rating_total)) %>% 
  slice_head(n=1)
```



